% CS 224R Milestone Report (5%) --- due 5/22/26.
\documentclass[10pt]{article}
\usepackage[margin=0.72in]{geometry}
\usepackage{times}
\usepackage[hidelinks]{hyperref}
\usepackage{booktabs}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.12em}
\emergencystretch=3em
\renewcommand{\arraystretch}{0.95}
\newcommand{\milesec}[1]{\par\vspace{0.35em}\noindent{\large\bfseries #1}\par\vspace{0.05em}}
\newcommand{\milesubsec}[1]{\par\vspace{0.20em}\noindent{\bfseries #1}\par\vspace{0.02em}}
\date{}

\begin{document}
\begin{center}
  {\large\bfseries Milestone Report\par}
  \vspace{2pt}
  {\Large\bfseries Self-Improving Agent Harnesses via Trace-Derived\\ RL Environments for Custom-Kernel Engineering\par}
  \vspace{6pt}
  {\normalsize Chandra Suda \quad Aaditya Nalawade\par}
  \vspace{1pt}
  {\small CS 224R: Deep Reinforcement Learning, Stanford \quad\textbar\quad May 23, 2026\par}
\end{center}
\vspace{-0.4em}
\linespread{0.94}\selectfont

\milesec{1.\ Experiments so far}
Our overarching objective is an ML-research agent that mines its own traces, diagnoses defects, synthesizes verifiable RL environments for those defects, and trains a lightweight LoRA policy that improves future harness behavior. For the milestone we narrowed the verifier domain to \emph{GPU kernel engineering} (\S2) and built \texttt{kernel-synth} (\url{https://github.com/chandrasuda/kernel-synth}): a trace-derived environment factory that shallow-clones a GitHub repo, walks its AST to enumerate every \texttt{nn.Module} subclass, and selects a diverse, non-obvious subset using either our custom tool-use agent harness or a deterministic AST heuristic favoring modules with unusual forward operations or imports of known custom-kernel libraries (Triton, FlashAttention, Mamba SSM). Each selected module is converted into a self-contained RL environment: a frozen reference, an auto-inferred input generator, an editable solution scaffold, a benchmarking harness, and metadata; a local FastAPI viewer browses the catalog. On top of this we wired a Harbor-style ATIF~v1.7 rollout layer that uses \texttt{torch.compile} as the reward ceiling and exposes Triton-only agent tools inside a workspace sandbox; baseline and \texttt{torch.compile} rollouts already validate against the schema.

\milesubsec{1.1\ Required experiment}
We ran the pipeline end-to-end in deterministic-heuristic mode on \textbf{five PyTorch repos chosen for mechanism diversity}:
\vspace{-0.6em}
\begin{center}
\footnotesize
\begin{tabular}{lrrrl}
\toprule
repo & py files & kept & avg novelty & top picks \\
\midrule
\texttt{state-spaces/mamba}        & 55 & 8 & 0.83          & Mamba2, Mamba3, MHA \\
\texttt{lucidrains/vit-pytorch}    & 76 & 8 & 0.92          & DSSA, NaViT, LookViT \\
\texttt{lucidrains/x-transformers} & 27 & 8 & 0.96          & CoPE, DataDependentAlibi, HyperConnection \\
\texttt{openai/whisper}            & 14 & 3 & \textbf{0.21} & (down-weighted: vanilla transformer) \\
\texttt{lucidrains/audiolm-pytorch}& 14 & 8 & 0.75          & Attend, SoundStream, HubertWithKmeans \\
\midrule
\textbf{total} & \textbf{186} & \textbf{35} & \textbf{0.81} & $\sim$60k LOC scanned \\
\bottomrule
\end{tabular}
\end{center}
\vspace{-0.5em}

\milesubsec{1.2\ Did results match expectations?}
\textbf{Largely yes.} We expected the heuristic to down-weight vanilla transformer code and up-weight unusual mechanisms; it did exactly that (Whisper $\Rightarrow$ avg 0.21; CoPE, DataDependentAlibi, Mamba2/3, DSSA $\Rightarrow$ near 1.0). This is the right inductive bias for our reward, since we don't want environments on modules where a fused kernel already exists (the speedup ceiling collapses). \textbf{Surprises.} The first converter na\"ively copied the source file's full import block, breaking many environments by pulling in unrelated dependencies. The fix (walk the class body for referenced names, keep only the imports actually used, transitively lift any helper definitions the class references) was straightforward but required a build-time correctness probe we hadn't planned for: trace-derived environments are only useful if the reference module actually runs.

\milesec{2.\ Changes to hypothesis / objective}
\textbf{The closed-loop hypothesis is unchanged}: agents improve faster when their RL training environments are mined from real trace/code distributions than when they are hand-designed. \textbf{What narrowed} is the verifier (``general ML-research harness defects'' $\to$ ``GPU kernel engineering''), because kernels admit a cheap, fully-automatable, \emph{dense} reward (numerical equivalence + wall-clock speedup vs \texttt{torch.compile}), which is exactly what makes long-horizon RL tractable; general ML-research decisions are subjective, slow to verify, and contaminated by API/cost noise, so they are better attacked once the loop is proven on the kernel sub-domain. \textbf{Preserved}: trace-derived envs, GRPO+LoRA training, evaluation vs base / SFT / prompt-only baselines, ablations over env-derivation vs synthetic. \textbf{New}: a concrete environment factory (\texttt{kernel-synth}), an external expert baseline (\texttt{torch.compile}) defining the reward ceiling, and Triton as a much smaller action space than the full agentic one (substantially easier to RL).

\milesec{3.\ Steps to completion}
\textbf{(1) Harden the rollout layer:} stress-test the per-environment benchmark pipeline, tighten the agent-mode tool-use loop, and improve the auto-inferred input generators so more environments run end-to-end out of the box. \textbf{(2)~Generate trajectories at scale:} run the latest frontier models and open-source coding LLMs over the 35 environments plus a held-out 5 repos to produce $N\!\approx\!200$ ATIF trajectories with rewards in $\{-0.1\}\cup[-0.2,1.5]$. \textbf{(3)~GRPO+LoRA training} on Qwen3-1.7B or SmolLM3-3B with reward $r=\mathrm{clip}((t_e-t_s)/(t_e-t_c),-0.2,1.5)$ ($t_e,t_s,t_c$~=~eager/solution/compile latency) gated by correctness; evaluate on held-out repos vs base, prompt-only self-reflection, and SFT. \textbf{(4)~Ablations:} (a)~heuristic vs agent env-selection, (b)~trace-derived envs vs a generic Triton-toy-task curriculum, (c)~\texttt{torch.compile} ceiling on/off. \textbf{(5)~Final eval:} best-of-$N$ solve rate, mean reward, fraction of held-out modules reaching $r\!\geq\!0.5$, and the $t_s/t_c$ distribution.

\milesubsec{3.1\ Scope realism and risks}
The remaining work is mostly trajectory generation + a small RL training loop on a $\leq$3B model, both of which we have run before. \textbf{Risks.} (i)~On CPU, \texttt{torch.compile} on small modules is often slower than eager, compressing the reward gap; mitigation is CUDA-aware benchmark presets that size shapes large enough for compilation to amortize. (ii)~Several environments still need hand-tuned input shapes; we will auto-detect the common cases (Embedding $\Rightarrow$ Long ids, Conv $\Rightarrow$ image tensors). (iii)~The environment distribution may concentrate around attention variants; we will pull held-out repos from disjoint mechanism families (state-space, audio codec, ViT, MoE) to widen coverage.

\milesec{4.\ Team contributions}
\textbf{Chandra Suda} (to date): kernel-synth data pipeline, AST heuristic for module selection, environment converter with import filtering and helper lifting, FastAPI viewer, Harbor-style ATIF rollout layer. \textbf{Aaditya Nalawade} (to date): related-work survey, reward-shape design, GRPO+LoRA training scaffold on Qwen3-1.7B, KernelBench-mini comparison baseline. \textbf{Planned split.} Chandra: agent-mode trajectory generation, held-out evaluation, env-selection ablations. Aaditya: training-loop integration, failure-mode analysis, KernelBench comparison. \textbf{Shared:} error analysis, final write-up, poster + presentation.

\end{document}
